---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r loadlibrary, echo=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(xtable)
library(lubridate)
```

```{r setoptions,echo=FALSE}
opts_chunk$set(echo=TRUE)
```


## Loading and preprocessing the data
1. Load the data

The raw file is in CSV format, so `read.csv` function will be used for loading the data.
```{r loaddata}
act <- read.csv("activity.csv")
```

2. Preprocess the data

The data has been in long-format, so I think that it isn't needed to preprocess the data any more.


## What is mean total number of steps taken per day?
1. Make the histogram

The histogram of the total number of steps taken each day is shown as:
```{r message=FALSE}
act.total <- act %>% filter(!is.na(steps)) %>% group_by(date) %>% summarise(total=sum(steps))
ph <- ggplot(act.total, aes(x=total)) + geom_histogram()
ph <- ph + labs(x="Total number of steps taken each day", y="Count")
ph
```

2. Calculate the mean and median

The **mean** and **median** total number of steps taken per day is shown as:
```{r results="asis"}
act.mm <- act %>% filter(!is.na(steps)) %>% group_by(date) %>% summarise(mean=mean(steps), median=median(steps))
xt <- xtable(act.mm)
print(xt, type="html")
```


## What is the average daily activity pattern?
1. Make the time series plot

The time series plot ot the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis), is shown as:
```{r}
act.ave <- act %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(ave=mean(steps))
pt <- ggplot(act.ave, aes(x=interval, y=ave)) + geom_line()
pt <- pt + labs(x="5-minute interval", y="Average number of steps")
pt
```

2. Find the maximum

The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps, is:
```{r message=FALSE, results="asis"}
act.ave.max <- act.ave %>% arrange(desc(ave)) %>% top_n(1)
xt <- xtable(act.ave.max)
print(xt, type="html")
```


## Imputing missing values
1. Calculate the total number of missing valuse
```{r}
act.na <- act %>% filter(is.na(steps)) %>% summarise(number=n())
```
The total number of missing values in the dataset (the total number of rows with NAs) is: `r act.na$number`.

2. Strategy for filling in missing values

Beacuse the missing values are always for a whole day, **the mean/median for that** will not be suitable. Finally, the **mean for that 5-minute interval** (rounded to the integer) will be used to fill in all of the missing values in the dataset.

3. Fill in the missing values

Create a new dataset that is equal to the original dataset but with the missing data filled in. The column named **step.new** is added.
```{r}
act.new <- act %>% group_by(interval) %>% mutate(steps.new=round(ifelse(is.na(steps), mean(steps, na.rm = TRUE), steps)))
```

4. Make the histogram and calculate the mean and median

* Make the histogram
```{r message=FALSE}
act.new.total <- act.new %>% group_by(date) %>% summarise(total=sum(steps.new))
ph2 <- ggplot(act.new.total, aes(x=total)) + geom_histogram()
ph2 <- ph2 + labs(x="Total number of steps taken each day", y="Count")
ph2
```

* Calculate the mean and median
```{r results="asis"}
act.new.mm <- act.new %>% group_by(date) %>% summarise(mean=mean(steps.new), median=median(steps.new))
xt <- xtable(act.new.mm)
print(xt, type="html")
```

* The impact of imputing missing data

After imputting the missing data, the histogram has changed slightly, for example, the max count at the y-axis and the shape at the middle of the x-axis (around 11000). Besides, which can be seen in the **mean** and **median** table, the number of rows (days) has been changed from `r nrow(act.total)` to `r nrow(act.new.total)`. The **mean** and **median** for days without missing data has not been changed, but those for days with missing data has been added.


## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable

The week begins from Monday(1) to Sunday(7), which can be seen at column named **week.num**. And **weekday** equals 1-5, **weekend** equals 6-7, which can be seen at column named **week**.
```{r}
act.week <- act.new %>% mutate(week.num=ifelse(wday(as.Date(date))==1,7,wday(as.Date(date))-1)) %>% mutate(week=as.factor(ifelse(week.num<=5,"weekday", "weekend")))
```

2. Make a panel plot
The panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis), is shown as:
```{r}
act.week.ave <- act.week %>% group_by(interval, week) %>% summarise(ave=mean(steps.new))
pt2 <- ggplot(act.week.ave, aes(x=interval, y=ave)) + geom_line()
pt2 <- pt2 + facet_wrap(~week, nrow=2)
pt2 <- pt2 + labs(x="5-minute interval", y="Average number of steps")
pt2
```
