---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r loadlibrary, echo=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(xtable)
library(lubridate)
```

```{r setoptions,echo=FALSE}
opts_chunk$set(echo=TRUE)
```


## Loading and preprocessing the data
1. Load the data

The raw file is in CSV format, so `read.csv` function will be used for loading the data.
```{r loaddata}
act <- read.csv("activity.csv")
```

2. Preprocess the data

The data has been in long-format, so I think that it isn't needed to preprocess the data any more.


## What is mean total number of steps taken per day?
1. Make the histogram

The histogram of the total number of steps taken each day is shown as:
```{r message=FALSE}
act.total <- act %>% filter(!is.na(steps)) %>% group_by(date) %>% summarise(total=sum(steps))
ph <- ggplot(act.total, aes(x=total)) + geom_histogram()
ph <- ph + labs(x="Total number of steps taken each day", y="Count")
ph
```

2. Calculate the mean and median

The **mean** and **median** total number of steps taken per day is shown as:
```{r results="asis"}
act.mm <- act %>% filter(!is.na(steps)) %>% group_by(date) %>% summarise(mean=mean(steps), median=median(steps))
xt <- xtable(act.mm)
print(xt, type="html")
```


## What is the average daily activity pattern?
1. Make the time series plot

The time series plot ot the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis), is shown as:
```{r}
act.ave <- act %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(ave=mean(steps))
pt <- ggplot(act.ave, aes(x=interval, y=ave)) + geom_line()
pt <- pt + labs(x="5-minute interval", y="Average number of steps")
pt
```

2. Find the maximum

The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps, is:
```{r message=FALSE, results="asis"}
act.ave.max <- act.ave %>% arrange(desc(ave)) %>% top_n(1)
xt <- xtable(act.ave.max)
print(xt, type="html")
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
